### 2020-01-30 wuhan.tex
minor tweaks

\newcommand{\swp}[1]{\comment{magenta}{SWP}{#1}}
\newcommand{\jd}[1]{\comment{magenta}{JD}{#1}}
\newcommand{\bmb}[1]{\comment{magenta}{BMB}{#1}}
\newcommand{\new}[1]{\textcolor{blue}{#1}}

@@

As the disease continues to spread, many researchers have published analyses of the outbreak, many focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
Estimating the basic reproductive number is of interest during an outbreak because it provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
However, it can be difficult to assess the validity of the estimates of \Ro (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can directly affect the estimates.

Here, we show that a wide range of approaches to estimating \Ro\ can be understood and compared in terms of estimates for three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$. We suggest that authors should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources.

We evaluate six disparate models posted online between January 24--26, 2020 that estimated \Ro\ for the 2019-nCoV outbreak. We use our framework to construct pooled estimates for the underlying parameters. We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval. We also use our framework to unravel which assumptions of these different models led to their different estimates and confidence intervals.

@@

\begin{table}[t]
\begin{center}
\footnotesize
\begin{tabular}{l|l|l|l}
 & Basic reproductive number & Mean generation time & CV$^2$ generation time \\
\hline
Study 1 & 2.5 (1.5--3.5)$^\ast$ & 8.4 days & unspecified$^\dagger$ \\
\hline
Study 2 & 2.92 (95\% CI: 2.28--3.67) & 8.4 days & 0.2 \\
\hline
Study 3 & 3.8 (95\% CI: 3.6--4.0) & 7.6 days & 0.5 \\
\hline
Study 4 & 2.2 (90\% CI: 1.4--3.8) & 7--14 days & 0.5\\
\hline
Study 5 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 days & 0.2\\
\hline
Study 6 & 2.0--3.1 & 6--10 days & 0 \\
\hline
\end{tabular}
\end{center}
\caption{
\textbf{Reported estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
Estimates of \Ro and their assumptions about the shape of the generation interval distributions were collected from 6 studies.
$^\ast$We treat these intervals as a 95\% confidence interval in our analysis.
$^\dagger$We assume $\kappa = 0.5$ in our analysis.
$^\ddagger$The authors presented \Ro estimates under different assumptions; we use their baseline scenario in our analysis.
}
\end{table}

>>>>

\newcommand{\swp}[1]{\comment{magenta}{SWP}{#1}}
\newcommand{\jd}[1]{\comment{magenta}{JD}{#1}}
\newcommand{\bmb}[1]{\comment{magenta}{BMB}{#1}}
\newcommand{\dc}[1]{\comment{Orange}{DC}{#1}}
\newcommand{\new}[1]{\textcolor{blue}{#1}}

@@

As the disease continues to spread, many researchers have published analyses of the outbreak, many focusing in particular on estimates of the basic reproductive number \Ro (i.e., the average number of secondary cases generated by a primary case in a fully susceptible population \citep{anderson1991infectious, diekmann1990definition}).
Estimating the basic reproductive number is of interest during an outbreak because it provides information about the level of intervention required to interrupt transmission \citep{anderson1991infectious}, and about the final size of the outbreak \citep{anderson1991infectious, ma2006generality}.
However, it can be difficult to assess the validity of the estimates of \Ro (as well as the associated degrees of uncertainty) when the estimation methods and their underlying assumptions vary widely, especially since these assumptions can directly affect the estimates.
\citep{Even if this is not directly addressed here, don’t you also want to talk about observation error/delays, which are clearly present with this outbreak?}

Here, we show that a wide range of approaches to estimating \Ro\ can be understood and compared in terms of estimates for three quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
The generation interval, which is defined as the time between when an individual becomes infected and when that individual infects another individual \citep{svensson2007note}, plays a key role in shaping the relationship between $r$ and \Ro\ \citep{wearing2005appropriate, roberts2007model, wallinga2007generation, park2019practical}.
We suggest that authors should make sure their assumptions about these three quantities are clear and reasonable, and that estimates of uncertainty should propagate error from all three sources.

We evaluate six disparate models posted online between January 24--26, 2020 that estimated \Ro\ for the 2019-nCoV outbreak. We use our framework to construct pooled estimates for the underlying parameters. We use these pooled estimates to illustrate the importance of propagating different sources of error, particularly uncertainty in both the growth rate and the generation interval. We also use our framework to unravel which assumptions of these different models led to their different estimates and confidence intervals.
\citep{Maybe add more explicitly somewhere in the introduction that inevitably, early in an epidemic, we have poor observations, hence large uncertainty, around case reporting (affecting r) and contact tracing (affecting g).}

@@

\begin{table}[t]
\begin{center}
\footnotesize
\begin{tabular}{l|l|l|l|c}
 & Basic reproductive number & Mean generation time & CV$^2$ generation time & Reference \\
\hline
Study 1 & 2.5 (1.5--3.5)$^\ast$ & 8.4 days & unspecified$^\dagger$ & \cite{imaincov} \\
\hline
Study 2 & 2.92 (95\% CI: 2.28--3.67) & 8.4 days & 0.2 & \cite{liuncov} \\
\hline
Study 3 & 3.8 (95\% CI: 3.6--4.0) & 7.6 days & 0.5 & \cite{readncov} \\
\hline
Study 4 & 2.2 (90\% CI: 1.4--3.8) & 7--14 days & 0.5 &  \\
\hline
Study 5 & 5.47 (95\% CI: 4.16--7.10)$^\ddagger$ & 7.6--8.4 days & 0.2\\
\hline
Study 6 & 2.0--3.1 & 6--10 days & 0 \\
\hline
\end{tabular}
\end{center}
\caption{
\textbf{Reported estimates of the basic reproductive number and the assumptions about the generation-interval distributions.}
Estimates of \Ro and their assumptions about the shape of the generation interval distributions were collected from 6 studies.
$^\ast$We treat these intervals as a 95\% confidence interval in our analysis.
$^\dagger$We assume $\kappa = 0.5$ in our analysis.
$^\ddagger$The authors presented \Ro estimates under different assumptions; we use their baseline scenario in our analysis.
}
\end{table}
